Handling String-Number Conversion in String Constraint Solving

### 2019-10-04 draft.tex
update

Based on the SFA encoding, we manage to translate the string constraint with string-number function to an equisatisfiable pure numerical constraints. Thus the problem is significantly simplified. We show that if we restrict the variable to arbitrary symbolic flat automata, one can convert the a string constraint with string-number function to a integer constraint consisting of both polynomials and exponentials (formula~\ref{formula:str_int}). To the best of our knowledge, the satisfiability prolem of integer constraints with a mix of polynomials and exponentials is still open.

@@

For the case that $n$ is a positive integer, the value of $n$ can be characterized by a constraint $n= (v_0\times 10+ v_1) \times (1+100 +100^2 + \ldots 100 ^{\#(v_0)-1})=  (v_0\times 10+ v_1) \times \frac{100^{\#(v_0)}}{100-1}$. 
The constraint uses $\#(v_0)$ to capture the total number of loop traversal.

@@

\[\begin{array}{lcl}
	y & = & (v_m^0\times 10^3 + v_m^1 \times 10^2 + v_m^2 \times 10 +v_m^3 )\times \frac{10000^{P(v_m^0)}-1}{10000-1}+v_{m-1}\times 10000^{P(v_m^0)}\\

	  &   &v_1\times10000^{P(v_m^0)} \times \ldots \times 1000^{P(v_1^0)}+\\
	  &   &(v_0^0\times 10^3 + v_0^1 \times 10^2 + v_0^2 \times 10 +v_0^3 )\times \frac{10000^{P(v_0^0)}-1}{10000-1} \times 10000^{P(v_m^0)} \times \ldots \times \\
	  &   &1000^{P(v_1^0)} \times 10
\end{array}\]

Recall that $P(x_1)$ is the number of occurrence of $v_1$. To solve the satisfiability of this formula, one need to solve an exponential constraint.

>>>>

Based on the SFA encoding, we manage to translate the string constraint with string-number function to an equisatisfiable pure numerical constraints. Thus the problem is significantly simplified. We show that if we restrict the variable to arbitrary symbolic flat automata, one can convert the a string constraint with string-number function to a integer constraint consisting of both polynomials and exponentials (formula~\ref{formula:str_int}). To the best of our knowledge, the satisfiability problem of integer constraints with a mix of polynomials and exponentials is still open.

@@

For the case that $n$ is a positive integer, the value of $n$ can be characterized by a constraint $n= (v_0\times 10+ v_1) \times (1+100 +100^2 + \ldots 100 ^{\#(v_0)-1})=  (v_0\times 10+ v_1) \times \frac{100^{\#(v_0)+1}-1}{100-1}$. 
The constraint uses $\#(v_0)$ to capture the total number of loop traversal.

However, this constraint contains an exponential component $100^{\#(v_0)}$. To the best of our knowledge, the satisfiability problem of integer constraints with a mix of polynomials and exponentials is still open. For a more general form, if we restrict the domain of $x$ to the SFA in Figure~\ref{fig:sfa_def}, for the case when $n$ is positive, we have the following relation.

\[\begin{array}{lcl}
y & = & (v_m^0\times 10^3 + v_m^1 \times 10^2 + v_m^2 \times 10 +v_m^3 )\times\\
&   & \frac{10000^{\#(v_m^0)+1}-1}{10000-1}+\\
&  & v_{m-1}\times 10000^{\#(v_m^0)+1}\\

&   &v_1\times10000^{\#(v_m^0)} \times \ldots \times 10000^{\#(v_1^0)}+\\
&   &(v_0^0\times 10^3 + v_0^1 \times 10^2 + v_0^2 \times 10 +v_0^3 )\times\\
&   &\frac{10000^{\#(v_0^0)}-1}{10000-1} \times 10000^{\#(v_m^0)} \times \ldots \times  10000^{\#(v_1^0)} \times 10
\end{array}\]

Recall that $\#(x_1)$ is the number of occurrence of $v_1$. To solve the satisfiability of this formula, one need to solve an exponential constraint.

For the particular case with only one integer variable, the satisfiability of integer constraints with a mix of polynomials and exponentials can be decided using the algorithm in~\cite{POPL2019}. However, the algorithm involves a quantifier elimination procedure, which is double-exponential to the length of the input formula and hence cannot handle large instances.

@@
